
\documentclass{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{enumerate}
\usepackage{mathtools}
\usepackage{xfrac}
\usepackage[lastexercise]{exercise}

\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\newcounter{Problem}
\newenvironment{Problem}{\begin{Exercise}[name={Problem},
                                          counter={Problem}]}
                        {\end{Exercise}}
\title{MATH 829 Homework \#3}
\date{February 4, 2015}
\author{Sam Boling}

\begin{document}

\begin{titlepage}
\maketitle
\end{titlepage}

\section{Textbook Problems}

\subsection*{Ch. II, \S 2.4}
\begin{itemize}
  \item[(a)]{
  }
  \item[(c)]{
  }
  \item[(f)]{
  }
  \item[(g)]{
  }
\end{itemize}

\subsection*{Ch. II \S 2.10}

\subsection*{Ch. III \S 6.1}
\begin{itemize}
  \item[(d)]{
  }
  \item[(e)]{
  }
\end{itemize}

\subsection*{Ch. III \S 6.2}
\begin{itemize}
  \item[(d)]{
  }
  \item[(e)]{
  }
\end{itemize}

\section{Additional Problems}

\begin{Problem}
  Let $\mathrm{Log} z$ be the principal logarithm function
  defined using $\mathrm{Arg} z \in (-\pi, \pi]$. Prove that
  $\mathrm{Log}(z)$ is holomorphic on
  $\mathbb{C} \backslash \{x \in \mathbb{R} : x \leq 0}$ and
  that $\frac{d}{dz} \mathrm{Log} z = \frac{1}{z}$.
  (Hint: Divide the domain into the union of the right, lower, and
  upper half-planes. Apply the Cauchy-Riemann equations.)
\end{Problem}

\begin{Problem}
  \begin{enumerate}[(i)]
    \item{
      Let $f$, $g$ be two entire functions such that
      $f^\prime(z) = g^\prime(z)$ for all $z$. Show that there is a
      constant $C \in \mathbb{C}$ such that $f = g + C$.
    }
    \item{
      Let $f$ be an entire function and $n \in \mathbb{N}$.
      Suppose $f^(n) = 0$. Show that $f$ is a polynomial
      with $\deg f \leq n - 1$.
    }
  \end{enumerate}
\end{Problem}

\begin{Problem}
  Let $f$ be an entire function. Suppose that $|f|$ is constant. Prove
  that $f$ is constant.
  (Hint: $|f| = C$ implies $u^2 + v^2 = C^2$. Take partial derivatives
  and apply the Cauchy-Riemann equations.)
\end{Problem}

\begin{Problem}
  Let $f$ be an entire function Suppose that $f^\prime(z) = f(z)$ for
  any $z \in \mathbb{C}$. Prove that $f(z) = C e^z$ for some
  $C \in \mathbb{C}$.
\end{Problem}

\begin{Problem}
  Let $\sum_{n=0}^\infty a_n z^n$ be a power series with radius $R$.
  \begin{enumerate}[(i)]
    \item{
      If the series diverges at $z = 3 - 4i$, what can you say about $R$?
    }
    \item{
      If the series converges at every $z \in D(0, 1)$, what can you say about
      $R$?
    }
    \item{
      What is the radius of $\sum_{n=0}^\infty a_n z^{2n}$?
    }
  \end{enumerate}
\end{Problem}

\end{document}
